Matching the Time of Emergence of Transboundary Fish Stocks to Lead Time for

### 2020-03-11 Scripts/Initial_Analysis.Rmd
Future proportion change by mid and end of the century

    if(c == 1){
      partial_ensamble <- world_data
    }else{
      spp_ensamble <- bind_rows(partial_ensamble,world_data)
    }

  }
  return(spp_ensamble)

@@

@@

>>>>

    if(c == 1){
      partial_ensamble <- world_data
    }else{
      spp_ensamble <- bind_rows(partial_ensamble,world_data)
    }
  }

  return(spp_ensamble)

@@

## Proportion approach

### Fun `EstPropChange`

```{r EstPropChange}

EstPropChange <- function(Spp,iyr,fyr,data_type,ensamble_list,Path,Neighbours){

  # Get model data from spps
  SppDist <- dbem_import(
    taxon_key = Spp,
    iyr <- iyr,
    fyr <- fyr,
    data_type <- "Abd",
    ensamble_list = ensamble_list,
    Path <- paste(data_path,"DBEM/",sep="")
  )

  # Result 1. Number of Countries that share the species

  #____________ Selecting only the transboundary nature of the species _________ #
  Trans_Spp <- SppDist %>%
    left_join(Neighbours,
              by= "index") %>% 
    filter(!is.na(value)) %>% 
    semi_join(Transboundary_spp, # filter only transboundary cases
              by = c("taxon_key","eez_name","eez_neighbour")
    )

  #____________ ESTIMATING DISTRIBUTION INDEX (TRESHOLD 3)_________ #
  # The number of species' cells present within each country's EEZ

  #Step 1.  Get EEZ id and Neighbour
  Neighbours_List <- Neighbours %>% 
    group_by(eez_name,eez_neighbour) %>% 
    summarise(n=n()) %>% 
    ungroup() %>% 
    select(-n)

  # Step 2. Determines the amount of grids present in each country
  Spp_Grid <- Trans_Spp %>% 
    group_by(taxon_key,
             eez_name,
             year,
             ensamble) %>% 
    summarise(n_spp_eez = length(unique(index))) %>% 
    left_join(Neighbours_List,
              by = "eez_name") %>% 
    filter(eez_name %in% Trans_Spp$eez_name, #Filter out unwanted Neighbours (those who don't have grids within but get included because they are Neighbours)
           eez_neighbour %in% Trans_Spp$eez_name)

  # Step 3. Sum total grids per Neighbours

  # Split dataframes to merge latter
  Territory_T <- Spp_Grid %>% 
    ungroup() %>% 
    select(
      taxon_key,
      Name=eez_name,
      n_spp_eez,
      year,
      ensamble
    )

  Neighbour_T <- Spp_Grid %>% 
    ungroup() %>% 
    select(
      taxon_key,
      n_spp_eez,
      Name=eez_neighbour,
      eez_name,
      year,
      ensamble
    )

  per_change_d <- full_join(Territory_T,
                            Neighbour_T, 
                            by = c("Name","taxon_key","year","ensamble")
  ) %>%
    rowwise() %>%
    mutate(Spp_Total = sum(n_spp_eez.x,n_spp_eez.y,na.rm=T)) %>% # Total gridcelles per Neighbours
    distinct() %>% # Removes false duplicates from `full_join()`
    rename(eez_name = Name,
           eez_neighbour =eez_name,
           n_spp_Country = n_spp_eez.x,
           n_spp_Neighbour = n_spp_eez.y) %>% 
    mutate(area_index = n_spp_Country/Spp_Total) %>% 
    select(taxon_key,year,ensamble,eez_name,eez_neighbour,area_index) %>% 
    mutate(
      time_step = ifelse(year < 2000,"historical",
                         ifelse(year >= 2045 & year <= 2064, "mid",
                                ifelse(year >= 2080, "n_end","NA")
                         )
      )
    ) %>%
    filter(time_step != "NA") %>% 
    # Time mean and sd
    ungroup() %>% 
    group_by(taxon_key,ensamble,eez_name,eez_neighbour,time_step) %>% 
    summarise_at(vars("area_index"),
                 c(temporal_mean=mean,
                   temporal_sd=sd)
    ) %>% 
    # Ensamble mean and sd
    group_by(taxon_key,eez_name,eez_neighbour,time_step) %>% 
    summarise_at(vars("temporal_mean"),
                 c(ensamble_mean=mean,
                   ensamble_sd=sd)
    ) %>% 
    mutate(ensamble_mean = ifelse(is.na(ensamble_mean),0,ensamble_mean)) %>% 
    select(-ensamble_sd) %>% 
    spread(time_step,ensamble_mean) %>% 
    mutate_at(vars(mid,n_end),
              funs(ifelse(historical == 0 & . == 0, 0, # If for no change than 0
                          ifelse(historical == 0 & . > 0, 100, # If changes from 0 to anything then 100
                                 ((.-historical)/historical)*100 # Estimate percentage
                                 # ((.-historical)/((.historical)/2))*100 # Estimate percentage
                          )
              )
              )
    )

  #### Statistical test ###

  stat_change_d <- full_join(Territory_T,
                             Neighbour_T, 
                             by = c("Name","taxon_key","year","ensamble")
  ) %>%
    rowwise() %>%
    mutate(Spp_Total = sum(n_spp_eez.x,n_spp_eez.y,na.rm=T)) %>% # Total gridcelles per Neighbours
    distinct() %>% # Removes false duplicates from `full_join()`
    rename(eez_name = Name,
           eez_neighbour =eez_name,
           n_spp_Country = n_spp_eez.x,
           n_spp_Neighbour = n_spp_eez.y) %>% 
    mutate(area_index = n_spp_Country/Spp_Total) %>% 
    select(taxon_key,year,ensamble,eez_name,eez_neighbour,area_index) %>% 
    mutate(
      time_step = ifelse(year < 2000,"historical",
                         ifelse(year >= 2045 & year <= 2064, "mid",
                                ifelse(year >= 2080, "n_end","NA")
                         )
      )
    )

  return(per_change_d)

### Fun `EstTransIndex`

```{r funEstTransIndex, eval= T ,echo = T,warning = F,message = F}

# Function to determine whether or not a species is transboundary

# Varibales needed
# Spp: Species to be analized, data with presence/absence per gridcell
# Index_EEZ: Reference list of all INDEX that fall within EEZs
# EEZ_Size: Number of grid-cells that each EEZ has
# Neighbours: Reference list of neighbouring countries

# Needs getSppDist

#### For testing 
Spp <- 600004  # Anchovie Equator, Chile and Peru
# Spp <- 600107 # Tuna Many RFMOs
# Spp <- 600004 # Argument 1 must have names (RFMO)
# Spp <- 600464 # Discrete species 
# Spp <- 600051 # One of Pierrs massive (wrong) discrete spp
# x <- c(600464, # Discreat spp in one nation
# 600059, # nowhere to be found spp
# 601352, # Discrete spp in two nations
# 600004) # transboundary spp
# Model <- "All"
# Coord <- CoorG
# 
# # For EEZ
# Neighbours <- Neighbours_Data
# Index_Code <- Index_Code

# __________________________________ #

# The function
EstTransIndex <- function(Spp,iyr,fyr,data_type,ensamble_list,Path,Neighbours){

  # Get model data from spps
  SppDist <- dbem_import(
    taxon_key = Spp,
    iyr <- iyr,
    fyr <- fyr,
    data_type <- "Abd",
    ensamble_list = ensamble_list,
    Path <- paste(data_path,"DBEM/",sep="")
  )

  # Result 1. Number of Countries that share the species

  #____________ Selecting only the transboundary nature of the species _________ #
  Trans_Spp <- SppDist %>%
    left_join(Neighbours,
              by= "index") %>% 
    filter(!is.na(value)) %>% 
    semi_join(Transboundary_spp, # filter only transboundary cases
              by = c("taxon_key","eez_name","eez_neighbour")
    )

  #____________ ESTIMATING DISTRIBUTION INDEX (TRESHOLD 3)_________ #
  # The number of species' cells present within each country's EEZ

  #Step 1.  Get EEZ id and Neighbour
  Neighbours_List <- Neighbours %>% 
    group_by(eez_name,eez_neighbour) %>% 
    summarise(n=n()) %>% 
    ungroup() %>% 
    select(-n)

  # Step 2. Determines the amount of grids present in each country
  Spp_Grid <- Trans_Spp %>% 
    group_by(taxon_key,
             eez_name,
             year,
             ensamble) %>% 
    summarise(n_spp_eez = length(unique(index))) %>% 
    left_join(Neighbours_List,
              by = "eez_name") %>% 
    filter(eez_name %in% Trans_Spp$eez_name, #Filter out unwanted Neighbours (those who don't have grids within but get included because they are Neighbours)
           eez_neighbour %in% Trans_Spp$eez_name)

  # Step 3. Sum total grids per Neighbours

      # Split dataframes to merge latter
    Territory_T <- Spp_Grid %>% 
      ungroup() %>% 
      select(
        taxon_key,
        Name=eez_name,
        n_spp_eez,
        year,
        ensamble
      )

    Neighbour_T <- Spp_Grid %>% 
      ungroup() %>% 
      select(
        taxon_key,
        n_spp_eez,
        Name=eez_neighbour,
        eez_name,
        year,
        ensamble
      )

    # Merge dataframes to get totals per Neighbourds
    area_index_d <- full_join(Territory_T,
                              Neighbour_T, 
                              by = c("Name","taxon_key","year","ensamble")
    ) %>%
      rowwise() %>%
      mutate(Spp_Total = sum(n_spp_eez.x,n_spp_eez.y,na.rm=T)) %>% # Total gridcelles per Neighbours
      distinct() %>% # Removes false duplicates from `full_join()`
      rename(eez_name = Name,
             eez_neighbour =eez_name,
             n_spp_Country = n_spp_eez.x,
             n_spp_Neighbour = n_spp_eez.y) %>% 
      mutate(area_index = n_spp_Country/Spp_Total) %>% 
      select(taxon_key,year,ensamble,eez_name,eez_neighbour,area_index) %>% 
      mutate(
        time_step = ifelse(year < 2000,"historical",
                           ifelse(year >= 2045 & year <= 2064, "mid",
                                  ifelse(year >= 2080, "n_end","NA")
                           )
        )
      ) %>%
      # Ensamble mean and sd
      ungroup() %>% 
      group_by(taxon_key,year,eez_name,eez_neighbour,time_step) %>% 
      summarise_at(vars("area_index"),
                   c(ensamble_mean=mean,
                     ensamble_sd=sd)
                   ) %>% 
      group_by(taxon_key,eez_name,eez_neighbour,time_step) %>% #Running mean
      mutate(RMean = rollmean(x = ensamble_mean,
                            5,
                            align = "right",
                            fill = ensamble_mean)
    )

    # Determine if changes
    natural_variation_d <- area_index_d %>% 
      ungroup() %>% #remove rowise
      filter(time_step == "historical") %>% 
      # Temporal mean
      group_by(taxon_key,eez_name,eez_neighbour) %>% 
      summarise_at(vars("ensamble_mean"),
                   c(hist_mean_index=mean,
                     hist_sd_index=sd)
      )

    ### Variation treshold
    variation_d <- area_index_d %>% 
      ungroup() %>% #remove rowise
      # Temporal mean
      group_by(taxon_key,time_step,eez_name,eez_neighbour) %>% 
      summarise_at(vars("RMean"),
                   c(mean_index=mean,
                     sd_index=sd)
      ) %>% 
      gather("variable","value",mean_index,sd_index) %>% 
      spread(time_step,value) %>% 
      mutate(larger_variation = ifelse(future > historical,"yes","no"))

    # William's index methodology
    trans_index_data <- area_index_d %>% 
      left_join(natural_variation_d,
                by = c("taxon_key","eez_name","eez_neighbour")) %>% 
      mutate(d_sd = ensamble_mean/sd_index) %>% 
      gather("variable","value",ensamble_mean:d_sd) %>% 
      ungroup() %>% 
      select(-time_step,-eez_neighbour) %>% 
      rowid_to_column() %>% 
      # spread(time_step,value) %>% 
      spread(eez_name,value) %>% 
      group_by(taxon_key,year,variable) %>% 
      summarise_at(vars("Chile","Ecuador","Peru"),
                   mean,na.rm=T)

    ggplot(area_index_d) +
      geom_ribbon(data=area_index_d,
                  aes(
                    x = as.numeric(year),
                    ymin=RMean-ensamble_sd,
                    ymax=RMean+ensamble_sd
                  ),
                  fill = "red",
                  alpha = 0.5
                  )+
      geom_line(
        aes(
          x = as.numeric(year),
          y = RMean
        )
      )  +
      facet_wrap(~eez_name + eez_neighbour,
                 scales = "free")

} # closes function

# __________________________________
# Test function
# __________________________________

@@

## Control Panel

```{r control_panel}

Spp_list <- c(600004,600005)
iyr <- 1951
fyr <- 2099
data_type <- "Catch"
ensamble_list = c(102,103)
Path <- paste(data_path,"DBEM/",sep="")
world_grid <- tibble(index = seq(1,259200,1))
Neighbours <- read_excel("Temporal_Data/Spatial/EEZ_Neighbour_List.xlsx") %>% 
  clean_names()
Transboundary_spp <- read.csv("~/GitHub/TransEmergence/Temporal_Data/Species/Transboundary_spp.csv") %>% clean_names()

## Main Routine

```{r main_routine, eval = T, echo = F}

####_______________________ ####
# Estimating proportion change
####_______________________ ####

suppressWarnings(
  r_proportion_change <- bind_rows(
    lapply(
      Spp_list, #c(600004,600005),
      EstPropChange,
      iyr <- iyr,
      fyr <- fyr,
      data_type <- "Abd",
      ensamble_list = ensamble_list,
      Path <- paste(data_path,"DBEM/",sep=""),
      Neighbours = Neighbours
    )
  )
)
```
